\authorsOfDoc{Marcel Rieser}
 
\bigskip

\begin{chapter-intro}
Starting a new scenario (our term for the application of MATSim to a 
region/area) can appear quite cumbersome at the first glace, as a lot of data
preparation may be required. This chapter gives first an overview of the input
data typically required for running a MATSim scenario, and then gives examples
how such data is generated for existing scenarios.
\end{chapter-intro}


\section{A Minimum Example}

In order to work, MATSim requires at least a network and a travel demand.
Alone with these two inputs and a configuration, MATSim can already be run.

The following section will describe these input data, along with other additional
and optional input data.

\todo{MR Expand this section} 



\section{Typical Input Data Sets}
MATSim uses multiple files to store the different types of data it uses for the
simulation. Tab.~\ref{tab:BuildingScenarios:InputDataSets} gives an
overview over files you may typically encounter when working with MATSim.

Not all files are always required. Very simple simulations can be run
only with a configuration file and the description of the network and
the population containing the agents' plans. For additional functionality, e.g.
for the simulation of public transport, additional files might be required.

\begin{table}[htp]
\begin{tabular}{ll}
\hline
{\tt config.xml}          & configuration options for MATSim \\
{\tt network.xml}         & description of the (road) network \\
{\tt population.xml}      & the travel demand, i.e. the list of agents
and their day plans
\\
{\tt facilities.xml}      & information about locations where
activities can be performed\\
{\tt transitSchedule.xml} & information about transit stop locations
and transit services\\
{\tt transitVehicles.xml} & description of the vehicles used for
public transport services\\
{\tt counts.xml}          & hourly volumes from real-world counting
stations for comparison\\
\hline
\end{tabular}
\caption{Files often used with MATSim}
\label{tab:BuildingScenarios:InputDataSets}
\end{table}

In the following, small examples of these files will be shown and the data they
contain discussed.

\begin{note}
Some of the files, especially {\tt population.xml}, but also {\tt network.xml} 
or {\tt facilities.xml}, might get quite large. To save space, MATSim supports
reading and writing the data in a compressed format. MATSim uses  
GZIP-compression for this. Thus, in many cases, the file names have the 
additional suffix {\tt .gz}, as in {\tt population.xml.gz}. MATSim automatically
detects if files are compressed or should be written compressed based on the 
filename.
\end{note}

\subsection{Configuration}

\begin{xml-file}[caption=An example of a config.xml,
label=lst:BuildingScenarios:configXml]
<?xml version="1.0" ?>
<!DOCTYPE config SYSTEM "http://www.matsim.org/files/dtd/config_v1.dtd">
<config>

	<module name="network">
		<param name="inputNetworkFile" value="example/network.xml" />
	</module>

	<module name="plans">
		<param name="inputPlansFile" value="example/population.xml.gz" />
	</module>

	<module name="controler">
		<param name="outputDirectory" value="./output/" />
		<param name="firstIteration" value="0" />
		<param name="lastIteration" value="10" />
	</module>
	
</config>
\end{xml-file}

The configuration file, often just referred to as \emph{config file}
or as \emph{config.xml}, builds the connection between the user and MATSim.
It contains a list of settings which influence how the simulation behaves.

All configuration parameters are simple pairs of a parameter name and a
parameter value. The parameters are grouped into logical groups. For example,
there is a group with settings related to the Controler like the number of
iterations, or there is another group with settings related to the simulation,
e.g. the end time of the simulation.
Listing~\ref{lst:BuildingScenarios:configXml} shows a very short example of a
configuration file which specifies the network and travel demand data to be used
along with some settings for the Controler.

The list of available parameters and valid parameter values may vary from
release to release. Although we try to keep this stable, due to changes in the
software, most notably by new features, settings may change. To get a list of
all available settings currently available, run the following command:
\begin{lstlisting}
java -cp matsim.jar org.matsim.run.CreateFullConfig fullConfig.xml
\end{lstlisting}
This command will create a new config file {\tt fullConfig.xml} which contains
the full list of available parameters along with their default values. This
makes it easy to see what settings are available. To use and modify certain
settings, the lines with the corresponding parameters can be copied to the
config file specific for the scenario to be simulated and the parameter values
be modified in that file.


\subsection{Network}

\begin{xml-file}[caption=An example of a network.xml,
label=lst:BuildingScenarios:networkXml]
<?xml version="1.0" encoding="utf-8"?>
<!DOCTYPE network SYSTEM "http://www.matsim.org/files/dtd/network_v1.dtd">

<network name="example network">
	<nodes>
		<node id="1" x="0.0" y="0.0"/>
		<node id="2" x="1000.0" y="0.0"/>
		<node id="3" x="1000.0" y="1000.0"/>
	</nodes>
	<links>
		<link id="1" from="1" to="2" length="3000.00" capacity="3600" 
		                           freespeed="27.78" permlanes="2" modes="car" />
		<link id="2" from="2" to="3" length="4000.00" capacity="1800" 
		                           freespeed="27.78" permlanes="1" modes="car" />
		<link id="3" from="3" to="2" length="4000.00" capacity="1800" 
		                           freespeed="27.78" permlanes="1" modes="car" />
		<link id="4" from="3" to="1" length="6000.00" capacity="3600" 
		                           freespeed="27.78" permlanes="2" modes="car" />
	</links>
</network>
\end{xml-file}

The network describes the infrastucture on which the agents (or the vehicles,
respectively), can move around. The network consists of \emph{nodes} and
\emph{links} (in graph theory, these are typically called \emph{vertices} and
\emph{edges}). Listing~\ref{lst:BuildingScenarios:networkXml} shows an example
of a simple description of a network in MATSim's XML data format.

Each element has an identifier \emph{id}. Nodes are described by an X and a Y
coordinate value. Links have more attributes: The \emph{from} and \emph{to}
attribute reference nodes and describe the geometry of the network. Additional
attributes describe the traffic-related aspects of the network:
\begin{itemize}\styleItemize
  \item the \emph{length} of the link, typically in meters (see
  Sec.~\ref{sec:BuildingScenarios:Units}).
  \item the \emph{flow capacity} of the link, i.e. the number of vehicles that
  can pass the link, typically in vehicles per hour.
  \item the \emph{freespeed} is the maximum speed at which vehicles are allowed
  to travel along the link, typically in meters per seconds.
  \item the \emph{number of lanes} ({\tt} permlanes) available in the direction
  specified by the {\tt from} and {\tt to} nodes.
  \item the list of \emph{modes} allowed on the link. This is a comma-separated
  list, e.g. {\tt modes="car,bike,taxi"}.
\end{itemize}
Note that all links are uni-directional. If a road can be travelled in both
directions, two links have to be defined with alternating {\tt to} and {\tt
from} attributes (see links with id {\tt 2} and {\tt 3} in the example given in
Listing~\ref{lst:BuildingScenarios:networkXml}).
Thus, the network can be seen as a directed graph.

\configOptionsNote{network}

\subsection{Demand}

The travel demand for MATSim is described by the agents' day plans. The full set
of agents is typically the \emph{population}, hence the filename {\tt
population.xml}. Alternatively, {\tt plans.xml} is also commonly used in MATSim,
as the population file essentially contains a list of day plans.

The population contains the data in a hierarchical structure, as shown in
Listing~\ref{lst:BuildingScenarios:populationXml}:
\begin{itemize}\styleItemize
  \item The population contains a list of persons.
  \item Each person contains a list of plans.
  \item Each plan contains a list of \emph{Activities} and \emph{Legs}.
\end{itemize}
Exactly one plan per person is marked as \emph{selected}. The selected plan of
each agent is the plan that gets executed by the mobility simulation. During the
replanning stage, a different plan might get marked as being selected. A plan
can contain a score as attribute. The score gets calculated and stored in the
plan during the scoring stage, after the plan was executed by the mobility
simulation.

The list of activities and legs in each plan describe the planned actions by
each agent. Activities have a type assigned and have---except for the last
activity in a day plan---an end time defined (There are some exceptions where
activities have a duration instead of an end time. Such activities are often
automatically generated by routing algorithms and are thus not described in
this guide). To describe the location where an activity takes place, the
activity is either assigned a coordinate by giving an x and y attribute value,
or has a link assigned which describes from which link the activity can be
reached. As the simulation requires the link attribute, the Controler calculates
the nearest link for a given coordinate in the case the attribute is missing and
only an x and y coordinate value is given or any activity.

\begin{xml-file}[caption=An example of a population.xml,
label=lst:BuildingScenarios:populationXml]
<?xml version="1.0" encoding="utf-8"?>
<!DOCTYPE population SYSTEM "http://www.matsim.org/files/dtd/population_v5.dtd">
<population>
	<person id="1">
		<plan selected="yes" score="93.2987721">
			<act type="home" link="1" end_time="07:16:23" />
			<leg mode="car">
				<route type="links">1 2 3</route>
			</leg>
			<act type="work" link="3" end_time="17:38:34" />
			<leg mode="car">
				<route type="links">3 1</route>
			</leg>
			<act type="home" link="1" />
		</plan>
	</person>
	<person id="2">
		<plan selected="yes" score="144.39002">
			\ldots
		</plan>
	</person>
</population>
\end{xml-file}

Legs describe how agents plan to travel from one location to the next one. Each
leg must have a transport mode assigned. Optionally, legs may have an
attribute {\tt trav\_time} which describes the expected travel time for this
leg. For a leg to be simulated, it must contain a route. The format of a
route depends on the mode of a leg. For car-legs, the route lists the links that
the agent has to travel along in the given order, while for transit-legs
information about the stop locations and expected transit services are stored.

An agent starts a leg directly after the previous activity (or leg) has ended.
Depending on the mode, the mobility simulation might handle the agent
differently. By default, car- and transit-legs are well-supported by the
mobility simulation. If the mobsim encounters a mode it does not know, it
defaults to \emph{teleportation}: In this case, the agent is removed from the
simulated reality, and after the leg's expected travel time has passed,
re-inserted at the agent's target location.

\bigskip

The population data format is one of the most central data structures in
MATSim and might be a bit overwhelming at first. Luckily, to get started, only a
small subset must be known of it.
Listing~\ref{lst:BuildingScenarios:minimalPopulationXml} shows how a minimal
population file could look like. Most notably, the following simplications can
be made:
\begin{itemize}\styleItemize
  \item Each person needs exactly one plan.
  \item The plan does not need to be selected or have a score.
  \item Activities can be located just by their coordinates.
  \item Activities should have a somewhat meaningful end-time.
  \item Legs only need a mode, but no routes.
\end{itemize}
When a simulation is started, MATSim's Controler will load such a file and then
automatically assign the nearest linnk to each activity and calculate a suitable
route for each leg. This makes it easy to get started quickly.

\begin{xml-file}[caption=Minimal population.xml required to start MATSim,
label=lst:BuildingScenarios:minimalPopulationXml]
<?xml version="1.0" encoding="utf-8"?>
<!DOCTYPE population SYSTEM "http://www.matsim.org/files/dtd/population_v5.dtd">
<population>
	<person id="1">
		<plan>
			<act type="home" x="5.0" y="8.0" end_time="08:00:00" />
			<leg mode="car">
			</leg>
			<act type="work" x="1500.0" y="890.0" end_time="17:30:00" />
			<leg mode="car">
			</leg>
			<act type="home" x="5.0" y="8.0" />
		</plan>
	</person>
	<person id="2">
		...
	</person>
</population>
\end{xml-file}

\configOptionsNote{plans}

\subsection{Public Transport}

To simulate public transport in MATSim, two additional input files are
necessary: One describes the schedule, i.e. that lines, their routes, which
stops they serve and the departure times. The other describes the vehicles which
serve the lines: are they big busses, small buses, trains or light rail
vehicles, and describes how many passengers each vehicle can transport.

The description of public transport vehicles itself can be split into two parts:
In a first part, vehicle types have to be described, specifying how many
passengers such a vehicle can transport (Note that the term "vehicle" can refer
to multiple vehicles in reality, e.g. a train with several wagons should be specified as
one long vehicle with a high number of seats). In the second part, actual
vehicles have to be listed. Each vehicle has an identifier and is of a previously
specified vehicle type. Listing~\ref{lst:BuildingScenarios:ptVehicles.xml} shows
an example of a such a file, describing one vehicle type and two vehicles of
that type.

\begin{xml-file}[caption=An example of transitVehicles.xml,
label=lst:BuildingScenarios:ptVehicles.xml]
<?xml version="1.0" encoding="UTF-8"?>
<vehicleDefinitions xmlns="http://www.matsim.org/files/dtd"
       xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance"
       xsi:schemaLocation="http://www.matsim.org/files/dtd 
                     http://www.matsim.org/files/dtd/vehicleDefinitions_v1.0.xsd">
<vehicleType id="1">
		<description>Small Train</description>
		<capacity>
			<seats persons="50"/>
			<standingRoom persons="30"/>
		</capacity>
		<length meter="50.0"/>
	</vehicleType>
	<vehicle id="tr_1" type="1"/>
	<vehicle id="tr_2" type="1"/>
</vehicleDefinitions>
\end{xml-file}

The public transport schedule, often referred to as \emph{transitSchedule.xml}
only, is a rather complex file. It contains information about stop facilities
(these can be bus stops, train stations or other stop locations) and transit
services.

In the first part, the stop facilities need to be defined, giving each one a
coordinate, an identifier and a reference to a link in the network. The stop can
only be served by vehicles driving on that specified link. Optionally, it is
possible to specify a name for the stop and if other vehicles are blocked
when a transit vehicle is waiting at a stop. This last attribute is useful to
model e.g. the difference of bus stops, where one bus stop has a bay, while at
another stop, the bus has to stop on the actual road.

After the stop facilities, the transit lines, their routes and schedules
are described. This is a hierarchical data structure: Each line can have one or
more \emph{routes}, each route has a \emph{route profile}, a \emph{network
route} and a list of \emph{departures}.
Listing~\ref{lst:BuildingScenarios:schedule.xml} has an example of a
minimalistic but complete transit schedule.

Each transit line must have a unique id. Each transit route has an id which must
be unique within that one line, so the same route id can be used with different
lines. The {\tt transportMode} describes on which links in the network the line
runs (Actually, this is currently not yet enforced. It would be possible to let
a bus run on train links in the simulation. It might be enforced in the future).

The {\tt routeProfile} describes the stops this route serves, while {\tt route}
itself describes the series of links in the network the transit vehicle's driver
has to drive along (thus often referred to as \emph{network route}. Note that
the complete route, i.e. all links the vehicle drives along, must be listed in
the route, and not only the ones where stops are located. All the specified
stops should occur along this route in the specified order. The time offsets
given for each stop in the routeProfile describe the relative time offset to an
actual departure time. If a bus is to depart at 7 o'clock in the morning, and
stop 2 has a departureOffset of {\tt 00:03:00}, this must be read that the bus
is expected to depart at 07:03 at the specific stop. All stops in the route
profile must have a departure offset definded, except the last one. All stops,
except the first one, can optionally have an arrival offset defined. This is
mostly useful for large trains that stop for several minutes at a station to
help the routing algorithm to find connecting services at the correct time,
namely the expected arrival time of the train.

As last part of the description of a transit
route, the list of departures should be given. Each departure has an id, which must be unique within the
route, and gives the departure time at the first stop of the specified
route profile. In addition, the departure specifies with which vehicle the
service should be run. This vehicle must be defined in the aforementioned
list of transit vehicles.

\begin{xml-file}[caption=An example of a schedule.xml,
label=lst:BuildingScenarios:schedule.xml]
<?xml version="1.0" encoding="UTF-8"?>
<!DOCTYPE transitSchedule SYSTEM "http://www.matsim.org/files/dtd/transitSchedule_v1.dtd">
<transitSchedule>
	<transitStops>
		<stopFacility id="1" x="990.0"  y="0.0"   name="Adorf"    
		                                          linkRefId="1" isBlocking="false"/>
		<stopFacility id="2" x="1100.0" y="980.0" name="Beweiler" 
		                                          linkRefId="2" isBlocking="true"/>
		<stopFacility id="3" x="0.0"    y="10.0"  name="Cestadt"  
		                                          linkRefId="3" isBlocking="false"/>
	</transitStops>
	<transitLine id="Blue Line">
		<transitRoute id="1">
			<description>Just a comment.</description>
			<transportMode>bus</transportMode>
			<routeProfile>
				<stop refId="1" departureOffset="00:00:00"/>
				<stop refId="2" arrivalOffset="00:02:30" departureOffset="00:03:00" 
				                                         awaitDeparture="true"/>
				<stop refId="3" arrivalOffset="00:05:00" awaitDeparture="true"/>
			</routeProfile>
			<route>
				<link refId="1"/>
				<link refId="2"/>
				<link refId="3"/>
			</route>
			<departures>
				<departure id="1" departureTime="07:00:00" vehicleRefId="12"/>
				<departure id="2" departureTime="07:05:00" vehicleRefId="23"/>
				<departure id="3" departureTime="07:10:00" vehicleRefId="34"/>
			</departures>
		</transitRoute>
	</transitLine>
</transitSchedule>
\end{xml-file}

Because of its complexity, transit schedules often contain little mistakes that
will return in an error when the simulation runs. Typical examples include that
the network route is missing a link, or that the network route does not pass at
all the defined stops in the right order. To make sure a schedule does not have
any such issues before the simulation is started, a special validation routine
is available:

\begin{verbatim}
java -Xmx512m -cp /path/to/matsim.jar 
      org.matsim.pt.utils.TransitScheduleValidator
      /path/to/transitSchedule.xml /path/to/network.xml
\end{verbatim}

If run, this validator will print out a list of errors or warnings, if any are
found, or show a message that the schedule appears to be valid.

\subsection{Counts}

MATSim provides funtionality to compare traffic volumes from your simulation to
real world values. The \emph{Counts} infrastructure allows to compare the 
traffic volumes on links on an hourly basis. Listing~\ref{lst:BuildingScenarios:counts.xml}
shows an example of a \emph{counts.xml} input file required to do traffic
count comparisons. It starts with a header containing general descriptive 
information about the counts, including a year to describe how current the 
data is. Next, for each link having real world counts data, the hourly volumes
can be specified. The network-link is referenced by the \emph{loc\_id} 
attribute, in the example, it's link {\tt 2}. The attribute \emph{cs\_id}
(counting station identifier) can be used to store an arbitrary description
of the counting station. Most often it is used to note the original real
word counting station to simplify future data comparison. The hourly volumes,
specified by the hour of the day (counting starts with hour "1") and its value,
are optional: That is, not for every hour a value must be given. If for a
counting station data is only available for certain hours of the day (e.g. only
during peak hours) it is possible to omit the other hours from the XML listing.  

\configOptionsNote{counts}

\begin{xml-file}[caption=An example of a counts.xml,
label=lst:BuildingScenarios:counts.xml]
<?xml version="1.0" encoding="UTF-8"?>
<counts xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance" 
        xsi:noNamespaceSchemaLocation="http://matsim.org/files/dtd/counts_v1.xsd" 
        name="test" desc="test counting stations" year="2014">
	<count loc_id="2" cs_id="005">
		<volume h="1" val="10.0"></volume>
		<volume h="2" val="1.0"></volume>
		<volume h="3" val="2.0"></volume>
		<volume h="4" val="3.0"></volume>
		<volume h="5" val="4.0"></volume>
		<volume h="6" val="5.0"></volume>
		<volume h="7" val="6.0"></volume>
		<volume h="8" val="7.0"></volume>
		<volume h="9" val="8.0"></volume>
		<volume h="10" val="9.0"></volume>
		<volume h="11" val="10.0"></volume>
		<volume h="12" val="11.0"></volume>
		<volume h="13" val="12.0"></volume>
		<volume h="14" val="13.0"></volume>
		<volume h="15" val="14.0"></volume>
		<volume h="16" val="15.0"></volume>
		<volume h="17" val="16.0"></volume>
		<volume h="18" val="17.0"></volume>
		<volume h="19" val="18.0"></volume>
		<volume h="20" val="19.0"></volume>
		<volume h="21" val="20.0"></volume>
		<volume h="22" val="21.0"></volume>
		<volume h="23" val="22.0"></volume>
		<volume h="24" val="23.0"></volume>
	</count>
</counts>
\end{xml-file}

\section{Units and Conventions Used}
\label{sec:BuildingScenarios:Units}

\subsection{Units}
MATSim tries to make as few assumptions about actual units as is possible,
but at some locations it cannot be done without any. In general, MATSim expects
similar values (e.g. all distances) to be in the same unit wherever they are
used. In the following, the most important (expected) units are listed in a 
short overview.

\paragraph{Distance} Distance units are most prominently used in links' length.
They should be specified in the same unit that the coordinate system uses. This
allows MATSim to use simple triangulation, e.g. with the nodes' coordinates,
to calculate beeline distances. As most of the typically used, projected 
coordinate systems (see Sec.~\ref{sec:BuildingScenarios:CoordinateSystems})
use meters as unit of distance, this is the most common used unit of distance
in MATSim.

\paragraph{Time} While MATSim supports an \emph{hour:minute:second} notation
in several places, internally it uses \emph{seconds} as the default time unit.
This implies that for example link speeds must be specified in distance per 
second, typically \emph{m/s}.

One noteable exception from this rule are scoring parameters, where MATSim
expects values per hour. This is due to the fact that most behavioral parameters
like value of time are typically estimated per minute or hour, and that the 
corresponding values for seconds are very small and thus errorprone to be
configured.


\subsection{Conventions}

MATSim makes heavy uses of identifiers, short \emph{Id}s. This Ids can be
arbitrary strings, with the following exceptions: Ids should not contain
any spaces (incl. tabs, new lines, etc) or commas, as those characters
are typically used for separating different Ids from each other in Id lists.


\section{Coordinate Systems}
\label{sec:BuildingScenarios:CoordinateSystems}

\subsection{Preparing Your Data in the Right Coordinate System}

In several input files, you need to specify coordinates, e.g. for the nodes of
the network. It is strongly suggested \emph{not} to use WGS84 coordinates (i.e.
GPS coordinates, or any other kind of spherical coordinates; coordinates ranging
from -180 to +180 in west-east direction, and from -90 to +90 in south-north
direction). MATSim needs to calculate distances between two points in several
places of the code. The calculation of distances between spheric coordinates is
very complex and potentially slow. Instead, MATSim uses the simple Pythagoras'
theorem, but this requires the coordinates to be in a Cartesian coordinate
system. Thus is is stronlgy advised to use a Cartesian coordinate system along
with MATSim, preferably one where the distance unit corresponds to one meter.

Many countries and regions have custom coordinate system defined, optimized for
usages in their apropriet areas. It might be best to ask some GIS specialists in
your region of interest what the most commonly used local coordinate system is
and use that as well for your data. 

If you don't have any clue about what coordinate system is used in your region,
it might be best to use the Universal Transverse Mercator coordinate system.
This coordinate system divides the world into multiple bands, each six degrees
width and separated into a northern and southern part, which it calls UTM zones
(see \url{http://en.wikipedia.org/wiki/UTM_zones#UTM_zone} for more details).
For each zone, an optimized coordinate system is defined. Choose the UTM zone
which covers your region (Wikipedia has a nice map showing the zones) and use
its coordinate system.

\subsection{Telling MATSim about Your Coordinate System}

In some places, MATSim requires to know which coordinate system your data is in.
You have multiple ways to specify the coordinate system you use. The easiest one
is to use the so-called ``EPSG codes''. Most of the commonly used coordinate
systems got standardized and numbered. The EPSG code uniquely identifies a
coordinate system and can be directly used by MATSim. As an alternative, MATSim
can also parse the description of a coordinate system in the so-called WKT
format. As the WKT format is much more error-prone it is suggested to use EPSG
codes whenever possible.

To find the correct EPSG code for your coordinate system (e.g. for one of the
UTM zones), the website \url{http://www.spatialreference.org} is of great use.
Search on this website for your coordinate system, e.g. for ``WGS84 / UTM Zone
8N'' (for the northern-hemisphere UTM Zone 8) to find a list of matching
coordinate systems along with their EPSG codes.


For some operations, MATSim must know the coordinate system your data is in.
Some analyses may create output to be visualized in Google Earth for example,
where the coordinates need to be converted back to WGS84. The coordinate system
used by your data can be specified in the config file:

\begin{lstlisting}{language=XML}
<module name="global">
  <param name="coordinateSystem" value="EPSG:32608" />
</module>
\end{lstlisting}

This allows MATSim to work with your coordinates and convert them whenever
needed.



\section{Reusing Existing Data}

Data from other modelling tools like PTV Visum or EMME can be partially reused.
Converting the network topology is often possible with little work. For Visum,
\href{http://matsim.org/javadoc/org/matsim/visum/package-summary.html}{some code
exists} that converts the network into MATSim's format. But note that this only
converts the network topology (location of nodes, and links with a limited set
of attributes). Due to the different working of other modelling tools (e.g. the 
use of speed-flow curves to describe links, with the BPR function being one of
the better knowns), attributes might be 
differently interpreted by MATSim (e.g. link capacities are hard limits in MATSim, 
vs. part of a function in other tools), leading to manual edits being needed 
after the initial conversion.

Reusing demand data is typically more complex. If you have the output of an 
agent-based and activity-based demand generation process, you might convert
the data agent by agent to the MATSim format. If only aggregated data
in the form of origin-destination matrices, or attraction matrices, is available,
a simple conversion to complex agent plans is not possible and the demand must be 
regenerated using other data sources.




%% 
%% ================
%% old stuff from here
%% 
%% 
%% Given these two data items, you can already start building your own
%% scenario.The "\href{http://www.matsim.org/docs/tutorials/learningIn3days}{Learning MATSim in 3 days}"-Tutorial gives you an introduction on how to build your own scenario.
%% 
%% 
%% \subsubsection{Programming}
%% 
%% In many cases, using pre-configured software is not possible because  there are just too many possibilities of how input could look like.  Although matsim is not there yet, these should be api-only use cases,  i.e. they should only use the "stable" api. Therefore, the  following are under the api-users section of the documentation:
%% \begin{itemize}
%% 	\item Additional information about network generation is \href{http://matsim.org/node/588}{here}.
%% 	\item additional information about initial demand generation is \href{http://matsim.org/node/340}{here}.
%% \end{itemize}
%% 
%% \subsubsection{Information concerning specific scenarios}
%% 
%% 
